# The beginning of the journey

## Day 1

### Five steps in omics data analysis

Generally speaking, I consider omics data analysis consisting of five steps:

(1) Data generation and collection. Why should we do an omics data analysis? Usually it's not for fun, though it indeed has a lot of fun.

Scientific studies usually aim to test a hypothesis, or generate new hypotheses. Usually it's a combination of both. That is to test a hypothesis, meanwhile to generate new hypotheses.

Depending on the hypothesis to test or generate, we need different data to answer.

(2) Data quality check and cleaning

(3) Data processing.

Sometimes, you need Supercomputerome, such as ([Computerome](https://www.computerome.dk/))

(4) Exploratory data analysis and modelling

(5) Visualization and reporting.

**Exercise:**

1.  Find out what is [Computerome](www.computerome.dk/)
2.  Which step(s) of omics data analysis is less familiar to you, and which may be a challenge for you?

### Getting started with R

#### R and RStudio

#### CRAN and Bioconductor

### Statistics

#### Why do we need statistics?

#### Classification and regression

## Day 2

### R: loading a package, and get help

library and package.

```{r day 2 load a library}
library(gprofiler2)
library(ggplot2)
library(dplyr)
```

### R: the basics

#### Arithmatic

```{r eval=FALSE}
2 + 3 * 5       # Note the order of operations.
log(10)         # Natural logarithm with base e
5^2             # 5 raised to the second power
3/2             # Division
sqrt(16)        # Square root
abs(3-7)        # Absolute value of 3-7
pi              # The number
exp(2)          # exponential function
28%%6           # Modulo
```

**Exercise:**

-   Calculate

    -   `r round(runif(1, min = -0, max = 200))` plus `r round(runif(1, min = -0, max = 200))`

    -   `r round(runif(1, min = -0, max = 200))` minus `r round(runif(1, min = -0, max = 200))`

    -   `r round(runif(1, min = -0, max = 200))` times `r round(runif(1, min = -0, max = 200))`

    -   the average of `r round(runif(1, min = -0, max = 200))` and `r round(runif(1, min = -0, max = 200))`

    -   the sum of `r round(runif(1, min = -0, max = 200))` and `r round(runif(1, min = -0, max = 200))`

    -   the difference between `r round(runif(1, min = -0, max = 200))` and `r round(runif(1, min = -0, max = 100))`

    -   the ratio between `r round(runif(1, min = -0, max = 200))` and `r round(runif(1, min = -0, max = 200))`

    -   `r round(runif(1, min = -0, max = 50))` multiplex `r round(runif(1, min = -0, max = 200))`

-   Execute `log(45)` and `log2(45)` and compare the difference. Note: you can get help using `help(log)` and `help(log2)`

#### Variable assignment

```{r}
# Assign the value 42 to x
x <- 42

# Print out the value of the variable x
x
```

```{r}
# Assign the value 5 to the variable my_apples
my_apples <- 5

# Print out the value of the variable my_apples
my_apples
```

```{r}
# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 6

# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
```

#### Variable type

```{r}
# Change my_numeric to be 42
my_numeric <- 42

# Change my_character to be "universe"
my_character <- "universe"

# Change my_logical to be FALSE
my_logical <- FALSE
```

```{r}
# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)

# Check class of my_logical
class(my_logical)
```

#### vector

```{r}
# Define the variable vegas
vegas <- "Go!"
```

```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")

# Complete the code for boolean_vector
boolean_vector <-c(TRUE,FALSE,TRUE)

# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Assign days as names of poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Assign days as names of roulette_vector
names(roulette_vector)<-c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

```{r}
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector
total_vector <- A_vector+B_vector
  
# Print out total_vector
total_vector
```
